---
title: "data_analysis"
author: "Lupe Antonio"
date: "2025-04-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
library(gtsummary)
```

# Analysis of Asthma-PASS Intervention 

```{r}
#loading data
data <- read_xlsx("Project 5/Proj 5 data.xlsx")

#clean up
data <- data %>%
  mutate(time = as.factor(time),
         group = as.factor(group),
         school = as.factor(school))
```

## Background:

Within inner city areas, like the Bronx, New York, asthma disproportionately affects low-income African-American and Hispanic children. To help manage asthma, physical activity plays an important role in children. It has been found to be associated with decreased severity in symptoms, reduced school absenteeism, and improved quality of life. Unfortunately, urban minority children with asthma often face barriers to physical activity on a personal, family, school, health care system, and community levels. Ideally, schools could optimize asthma management and promote physical activity. However, often times, they do not faciliate appropriate asthma management and related exercise and may even discourage it. Consequently, there is a need for comprehensive school-based interventions (in collaboration with communities) to help reduce asthma morbidity and promote physical activity in urban minority children with persistent asthma.

Therefore, investigators proposed an intervention program in collaboration with primary care physicians (PCPs) to promote availability of guidline-based preventive medications and rescue medication at schools. The community health workers (CHW) will deliver education to children, caregivers, and school personnel encouraging physical activity. 

A pilot cluster-RCT, where the unit of randomization is school and not students, of Asthma-PASS in four Bronx elementary schools was conducted. A total of 108 asthmatic children were recruited.  



## Objective 1:

Investigators are interested in whether children in schools receiving Asthma-PASS intervention may experience a greater improvement in number of SFD (Symptom Free Days) at 6 or 12 months follow up compared to children in the AM comparison group.



## Exploratory Data Analysis:

Asthmatic children were recruited from four Bronx elementary schools among 104 Bronx elementary schools that expressed interested in the study. For students in Bronx and other NYC schools, between 89%-99% were eligible for free lunch (a marker of poverty), 48%-52% were male, and 62%-78% were Hispanics and 20-38% were African Americans. 

A total of 108 children between the ages of 4-11 with physician-diagnosed persistent or uncontrolled asthma attending kindergarten to 5th grade (from the 4 schools selected) were recruited. The four elementary schools were randomly assigned to Asthma-Pass intervention group (2 schools) or AM comparison group (2 schools), where participants followed the routine provided by each school (the standard care). The particpants were followed at 6 and 12 months post baseline. 

```{r, table 1}
data %>%
  distinct(ID, .keep_all = TRUE) %>%
  select(-ID, -time) %>%
  tbl_summary(by = group,
              statistic = list(all_continuous() ~ '{mean}({sd})',
                               all_categorical() ~ '{n} ({p}%)'),
              digits = all_continuous() ~ 3,
              label = list(group ~ 'Group')) %>%
  modify_spanning_header(c('stat_1', 'stat_2') ~ '**Intervention Group**') %>%
  modify_header(label = '**Variable**',
                stat_1 = paste('**Control**, N = 49'),
                stat_2 = paste('**Intervention**, N = 59')) %>%
  bold_labels() %>%
  modify_caption('Table 1. Summary statistics')
```


```{r, vals for table 1}
#values to modify table:

#avg SFD at baseline,6,12 per group:
data %>%
  filter(time == 3) %>%
  distinct(ID, .keep_all = TRUE) %>%
  group_by(group) %>%
  summarise(avg_sfd = mean(SFD, na.rm = TRUE)) 

#0 (9.60), 1(8.95) @ t = 1
#0 (11), 1(11.9) @ t = 2
#0 (10.9), 1(11.6) @ t = 3

#avg SFD per group at each time point:
data %>%
  group_by(group, time) %>%
  summarise(avg_sfd = mean(SFD, na.rm = TRUE))

#overall avg SFD per group
data %>%
  group_by(group) %>%
  summarise(avg_sfd = mean(SFD, na.rm = TRUE))
#0 (10.5), 1 (10.8)


data %>%
  filter(time == 3 & group == 1) %>%
  distinct(ID, .keep_all = TRUE) %>%
  drop_na()
#n = 48 in control at time 1
#n = 46 '' @ t = 2
#n = 43 '' @ t = 3

#n = 57 in interv at time 1
#n = 52 '' @ t = 2
#n = 50 '' @ t = 3


#to see how many participants had at least one missing value of SFD at any time point:
data %>%
  group_by(ID) %>%
  filter(any(is.na(SFD))) %>%
  distinct(ID)

# 1 1012_1 #group 1, time 2, 3
# 2 1021_1 #group 1, time 1, 2, 3
# 3 1026_1 #group 1, time 2, 3
# 4 1027_1 #group 1, time 3
# 5 1028_1 #group 1, time 3
# 6 2011_1 #group 0, time 2, 3
# 7 2015_1 #group 0, time 2
# 8 3001_1 #group 1, time 2, 3
# 9 3007_1 #group 1, time 1, 2, 3
#10 3009_1 #group 1, time 3
#11 3016_1 #group 1, time 2, 3
#12 3023_1 #group 1, time 2
#13 4009_1 #group 0, time 3
#14 4011_1 #group 0, time 3
#15 4012_1 #group 0, time 1, 2, 3
#16 4025_1 #group 0, time 3
#17 4026_1 #group 0, time 3

#group 0; time 1=1, time 2=3, time 3=6
#group 1; time 1=2, time 2=7, time 3=9

#total of 17 ppl with missing values.
```


```{r, not ideal plot}
#SFDs over time by group:
data %>% group_by(time, group) %>%
  summarise(sum_sfd = sum(SFD, na.rm = TRUE)) %>%
  ggplot(aes(x = time, y = sum_sfd, col = group)) +
  geom_point() +
  geom_line(aes(group = group)) +
  xlab('Follow-Up Times') + 
  ylab('# of symptom free days') + 
  theme_classic()

```

```{r, avg SFD over time per group}
sfd_data <- data %>%
  group_by(group, time) %>%
  summarise(avg_sfd = mean(SFD, na.rm = TRUE),
            sd_sfd = sd(SFD, na.rm = TRUE),
            n = sum(!is.na(SFD)),
            se_sfd = sd_sfd/sqrt(n)) %>%
  ungroup()

ggplot(aes(x = time, y = avg_sfd, col = group, group = group), data = sfd_data) +
  geom_point() + 
  geom_line() +
  xlab('Follow-up Times') + 
  ylab('Average Number of SFDs (out of 14)') + 
  theme_classic() +
  scale_color_brewer(palette="Set1")
```


## Binomial Mixed Effects model:

Justification: Our outcome of interest is SFD which has a max value of 14 days `max(data$SFD, na.rm = TRUE)`. We also have participant variation and school variation. 

```{r}
#SFD response ~ Binomial(14, p)
#since each day yes or no have symptom, so sum up the # of symptom free days


library(lme4)

mod1 <- glmer(cbind(SFD, 14-SFD) ~ group*time + (1 | ID/school),
              data = data,
              family = 'binomial')

summary(mod1)
#AIC = 1896.3
#time2, time3, g1:t2, g1:t3 are significant


mod1_noint <- glmer(cbind(SFD, 14-SFD) ~ group + time + (1 | ID/school),
                    data = data,
                    family = 'binomial')

summary(mod1_noint)
#AIC = 1915.8
#t2, t3 signf


anova(mod1_noint, mod1)
#null: no interaction model is better/reduced model better
#reject null -- low p-val, model with interaction is significant.
#interaction improves model fit
#baseline as covariate
```

The interaction between time and group is significant. This indicates that the change in SFDs over time differs by group. Since the coefficients are positive (for interaction terms), we can conclude that the intervention group experienced greater improvement over time compared to the control group. Meaning, we see an increase in SFDs over time compared to the control group. 

Coefficient interpretations:
currently coefficients are log-odds of having a symptom free day

time2: the control group improves from baseline to 6 months significantly.
       The odds of SFDs at 6 months are about 1.99 more compared to baseline for the control group.
       
time3: The odds of SFDs at 12 months are about 1.87 higher compared to baseline for the control group.

g1:t2: For those in the intervention group, the odds of SFDs at 6 months increases by 5.07 (1.99*2.55) compared to the control group at baseline?
Meaning, those in the intervention group experienced much greater improvement in odds of SFDs at 6 months compared to the control group. 

g1:t3: For those in the intervention group, the odds of SFDs at 12 months increases by 4.06 (1.87*2.17) compared to baselines.
Meaning, similar to above. Significant improvement of SFDs over time.

```{r}
#plotting SFD distribution:
#overall
data %>%
  ggplot(aes(x = SFD)) + 
  geom_histogram()

#at time 2:
data %>%
  filter(time == 2) %>%
  ggplot(aes(x = SFD)) + 
  geom_histogram()

#at time 3:
data %>%
  filter(time == 3) %>%
  ggplot(aes(x = SFD)) + 
  geom_histogram()

#this can go well in EDA to show zero-inflation
```


```{r}
#diagnostic checking for mod above:
library(DHARMa)

#residual plots
plot(simulateResiduals(fittedModel = mod1, n = 1000))

#overdispersion:
testDispersion(simulateResiduals(fittedModel = mod1, n = 1000))

#zero-inflation:
testZeroInflation(simulateResiduals(fittedModel = mod1, n = 1000))
```


***effect size, treatment effect***

```{r, other option}
library(glmmTMB) #supports zero-inflation, dispersion models

mod2 <- glmmTMB(cbind(SFD, 14-SFD) ~ group*time + (1 | ID/school),
                data = data,
                family = 'binomial')

summary(mod2)

mod2_noint <- glmmTMB(cbind(SFD, 14-SFD) ~ group + time + (1 | ID/school),
                      data = data,
                      family = 'binomial')

summary(mod2_noint)
#AIC = 1915.8

anova(mod2_noint, mod2)
```


```{r, binomial zero inflation mod}
mod3 <- glmmTMB(cbind(SFD, 14-SFD) ~ group*time + (1 | ID/school),
                zi = ~1,
                data = data,
                family = 'binomial')

summary(mod3)
#AIC = 1302.5 (best)
#time2, time3 significant
#Zero-Inflated Beta Binomial
```

```{r}
#baseline data only:
baseline_data <- data %>%
  filter(time == 1) %>%
  select(ID, baseline = SFD)

#6,12 month data only:
fu_data <- data %>%
  filter(time == 2 | time == 3)

bbz_data <- fu_data %>%
  left_join(baseline_data, by = 'ID')
  
  
#this models handles overdispersion + zero-inflation.

mod4 <- glmmTMB(cbind(SFD, 14-SFD) ~ baseline + group*time + (1 | ID/school),
                zi = ~1,
                data = bbz_data,
                family = 'betabinomial')

summary(mod4)
#AIC = 708.4
```

