---
title: "Project 2 - Survival"
author: "Lupe Antonio"
date: "2025-03-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(ggplot2)
library(DescTools)

library(flexsurv)
library(survival)
library(MASS)
library(survminer)
```



##loading & cleaning data:
```{r}
menopause_dat <- read.delim("Project 2 - Survival/Menopause.dat", header = FALSE)

menopause <- menopause_dat %>%
  rename(id = V1,
         intake_age = V2,
         menopause_age = V3,
         menopause = V4,
         race = V5,
         education = V6) %>%
  mutate(menopause_time = menopause_age - intake_age,
         race = as.factor(race),
         education = as.factor(education),
         menopause = as.integer(menopause)) 


#description of variables:
#menopause = 1 (observed to experience menopause), 
#            0 (censored at menopause age)

#race = 0 (White, non-hispanic),
#       1 (Black, non-hispanic),
#       2 (Other)

#education = 0 (post-grad),
#            1 (college grad),
#            2 (some college),
#            3 (high school or less)


#menopause time = the duration of time in study at which the subject experienced menopause
```


#Part A

##IA. Estimate median menopause time for all subjects, disregarding covariates. 

note: suppose that menopause time distribution is approx. exponential. 


##exploring distribution of menopause time:
```{r}
ggplot(data = menopause, aes(x = menopause_time)) +
  geom_histogram()
```


#computing median menopause time:
```{r}
median(menopause$menopause_time)
```

4.0424

#computing CI:
```{r}
MedianCI(menopause$menopause_time, conf.level = 0.95)
```

  median   lwr.ci   upr.ci 
4.042437 3.874059 4.254620 
attr(,"conf.level")
[1] 0.954713



actual estimation of menopause time:
```{r}
lambda_hat <- 1 / (sum(menopause$menopause_time)/sum(menopause$menopause==1))  # MLE for exponential rate
median_exp <- log(2) / lambda_hat
cat("Estimated Median (Exponential Model):", median_exp, "years.\n")
```




##IB. Nonparametric estimate of survival fn for menopause time (using KM)


```{r}
#KM survival fn

km_mt <- survfit(Surv(menopause_time, menopause) ~ 1, data = menopause)


#plotting K-M curve
plot(km_mt, mark.time = TRUE, xlab = "Years", ylab = "Survival Probability",
     main = "Menopause time K-M curve", cex.main = 1.5)

#another plot possibility
ggsurvplot(km_mt, data = menopause, conf.int = TRUE, 
           ggtheme = theme_classic(), xlab="Time to Menopause (years)", 
           ylab="Survival Probability")



#plotting cumulative hazard fn
plot(km_mt, fun = 'cumhaz')


#estimating cumulative hazard rates
cbind(km_mt$time, -log(km_mt$surv), cumsum(km_mt$n.event/km_mt$n.risk))
#time , K-M estim, Nelson-Aalen Estimator
#cumulative

#surival rate @ given time, w CI
summary(km_mt, time = c(0, 1, 2, 3, 4, 5))
#note: n.event is the cumulative num of events since prev time

summary(km_mt, censored = TRUE)
#not specifying time, n.event = # event @ each time point


#somehow median surival time according to km-curve is not outputting
print(km_mt)
```


```{r}
#table of estimates for km curve
km_table <- data.frame(time = km_mt$time,
                  n.risk = km_mt$n.risk,
                  n.event = km_mt$n.event,
                  n.censor = km_mt$n.censor,
                  surv = km_mt$surv,
                  upper = km_mt$upper,
                  lower = km_mt$lower)

print(km_table)
```

possibly add slide with survival function?


II. 

fitting cox model.

```{r}
cox_model <- coxph(Surv(menopause_time, menopause) ~ race + education + intake_age, data = menopause)


summary(cox_model)

cox_model2 <- coxph(Surv(menopause_time, menopause) ~ intake_age, data = menopause)


anova(cox_model, cox_model2) #doesnt matter whether include or not. #for better interpretation include race + education #actually big model is better
```

checking assumptions: assume proportion is constant overtime


The Schoenfeld residuals are calculated for all covariates for each individual experiencing an event at a given time. Those are the differences between that individual's covariate values at the event time and the corresponding risk-weighted average of covariate values among all those then at risk. The word "residual" thus makes sense, as it's the difference between an observed covariate value and what you might have expected based on all those at risk at that time.

```{r}
# Check the proportional hazards assumption
cox.zph_test = cox.zph(cox_model)

# Plot Schoenfeld residuals to visualize proportional hazards assumption
ggcoxzph(cox.zph_test)

```





III. 

exponential distribution: 
```{r}
lambda_hat <- 1 / (sum(menopause$menopause_age)/sum(menopause$menopause==1))  # MLE for exponential rate
median_exp <- log(2) / lambda_hat
cat("Estimated Median (Exponential Model):", median_exp, "years.\n")

#not realistic and KM-estimator is better outcome. 
```

S(t) = same as before

```{r}
km_fit_age <- survfit(Surv(intake_age, menopause_age, menopause) ~ 1, data = menopause)
```

```{r}
km_age_table <- data.frame(age = km_fit_age$time,
                           n.risk = km_fit_age$n.risk,
                           n.event = km_fit_age$n.event,
                           n.censor = km_fit_age$n.censor,
                           survival = km_fit_age$surv, 
                           lower = km_fit_age$lower, 
                           upper = km_fit_age$upper)

print(km_age_table)
```

```{r}
cat("Estimated median menopause age (Kaplan-Meier):", summary(km_fit_age)$table["median"], "years.\n")
```

survfit2
ggsurvfit




146.98 232.98


```{r}
cox_model_age <- coxph(Surv(menopause_age, menopause) ~ race + education, data = menopause)

cox.zph_test = cox.zph(cox_model_age)

ggcoxzph(cox.zph_test)
```

