---
title: "data analysis"
author: "Lupe Antonio"
date: "2025-04-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
library(lme4)
library(naniar)
library(mice)
library(gtsummary)
library(ggplot2)
library(corrplot)
library(rstpm2)
```

# Background



# A

Write statistical analysis plan, includes (1) randomization procedure, (2) analytic approach for primary analysis, (3) sample size calculation.

- Relative risk: [ppl got sick & vaccine/ppl got vaccine] / [ppl got sick & no vaccine/ppl no vaccine]



# B

```{r}
#loading data:

demographics <- read_xlsx("Project 4/Q2b_BL.xlsx")

followup <- read_xlsx("Project 4/Q2b.xlsx")
```

## Objective: 

Determine whether the vaccine group has greater odds of having severe adverse effects (SAE) at any of the three assessment time points.


SAE - binary

"SAE was defined as any untoward medical occurrence that was life-threatening or required inpatient hospitalization."

```{r}
#data cleaning + merging

followup_wide <- followup %>%
  pivot_wider(names_from = TIME,
              values_from = SAE)


data_wide <- merge(demographics, followup_wide, by = 'ID') %>%
  mutate(SITE = as.factor(SITE),
         SEX = as.factor(SEX),
         )
  

#final:
data <- data_wide %>%
  pivot_longer(cols = '1':'3',
               names_to = 'TIME',
               values_to = 'SAE') %>%
  mutate(SITE = as.factor(SITE),
         SEX = as.factor(SEX),
         TIME = as.factor(TIME),
         SAE = as.factor(SAE),
         GROUP = as.factor(GROUP))

#12298 SAE NA's
```


```{r, table 1}
data %>%
  distinct(ID, .keep_all = TRUE) %>%
  select(-ID, -missing) %>%
  #mutate(
  #  Gender = case_when(Gender == 'M' ~ 'Male',
  #                     Gender == 'F' ~ 'Female'),
  #  
  #  Treatment_Group = case_when(Treatment_Group == 'A' ~ 'Placebo',
  #                              Treatment_Group == 'B' ~ 'Low dose',
  #                              Treatment_Group == 'C' ~ 'High dose')) %>%
  tbl_summary(by = GROUP,
              statistic = list(all_continuous() ~ '{mean}({sd})',
                               all_categorical() ~ '{n} ({p}%)'),
              digits = all_continuous() ~ 3,
              label = list(GROUP ~ 'Group')) %>%
  #modify_spanning_header(c('stat_1', 'stat_2', 'stat_3') ~ '**Day**') %>%
  modify_header(label = '**Variable**') %>%
  bold_labels() %>%
  modify_caption('Table 1. Summary statistics')
```


```{r}
data %>%
  distinct(ID, .keep_all = TRUE) %>%
  filter(GROUP == 0 & TIME == 1) %>%
  drop_na()

data %>%
  filter(GROUP == 0 & TIME == 2) %>%
  distinct(ID, .keep_all = TRUE) %>%
  drop_na()
```


```{r}
ggplot(data = data, aes(x = as.factor(missing))) +
  geom_bar()
```


```{r, checking multicollinearity}
cor_data <- data_wide %>%
  select(-ID, -SITE) %>%
  mutate(SEX = as.numeric(SEX))
#%>%
 # mutate(GROUP = as.factor(GROUP))

corrplot(cor(cor_data, use='complete.obs'), addCoef.col = "red", type = "upper",tl.col="black", tl.srt=20)
```


```{r, Little's Test}
mcar_test(data)

#H_0: data MCAR (missing completely at random)

#high chi-squared stat & low pval == reject null, not MCAR
```


```{r, checking MAR}
data <- data %>%
  mutate(missing = is.na(SAE),
         missing = case_when(
           missing == 'TRUE' ~ 1,
           missing == 'FALSE' ~ 0))

missing_mod <- glm(missing ~ GROUP + TIME + SEX + AGE + SITE,
                   data = data,
                   family = 'binomial')

summary(missing_mod)

#group, time seem to predict missingness well. (missingness predicted from obs data)
#suggests MAR
#thus, logistic mixed effects mod is robust under MAR. no need to impute.
```


```{r}
gg_miss_fct(data, fct = GROUP)

vis_miss(data)
```


```{r}
md.pattern(data)
```



```{r, eval = FALSE}
mod1 <- glmer(SAE ~ GROUP + TIME + SITE + SEX + AGE + (1 | ID),
              data = data,
              family = 'binomial')


summary(mod1)

#not running; probs bc site has 100 levels
```


```{r, no interaction mod}
mod2 <- glmer(SAE ~ GROUP + TIME + SEX + AGE + (1 | ID),
              data = data,
              family = 'binomial')

summary(mod2)

#AIC = 2615.2 (LOWER = BETTER)
#time levels not signif.
#age is only signif at .05
```


```{r, interaction group*time mod}
mod3 <- glmer(SAE ~ GROUP*TIME + SEX + AGE + (1 | ID),
              data = data,
              family = 'binomial')

summary(mod3)

#AIC = 2617.8
#age only signif at .05
```


```{r, final mod}
mod4 <- glmer(SAE ~ GROUP*TIME + SEX + AGE + (1 | ID/SITE),
              data = data,
              family = 'binomial')

summary(mod4)
#AIC: 2619.8
```

```{r, normality of random effects}
rand_effs <- ranef(mod4, condVar = TRUE)

#plotting
qqnorm(rand_effs$`SITE:ID`[[1]], main = 'Q-Q plot of RE (Patient within Site)')

qqnorm(rand_effs$ID[[1]])
```


```{r}
mod4_2 <- glmer(SAE ~ GROUP + TIME + SEX + AGE + (1 | ID/SITE),
              data = data,
              family = 'binomial')

summary(mod4_2)

anova(mod4, mod4_2)
#null: no interaction model is better/reduced model better
#fail to reject, then interaction is not significant,
#thus, also supports that SAE odds to not change/vary by group or over time

#this is the correct conclusion right
```


```{r}
mod5 <- glmer(SAE ~ GROUP*TIME + SEX + AGE + (1 | SITE),
              data = data,
              family = 'binomial')

summary(mod5)
#random effect must be linear?
#AIC: 2616.3

rand_effs <- ranef(mod5, condVar = TRUE)

qqnorm(rand_effs$SITE[[1]])
```

ID:SITE


#pattern of missing; MAR -- use MIXED EFFECTS MODEL

ANOVA - time points, non-signifc. interaction use,


#consider sensitivity analysis:

- compare completers vs non, to see if there is much of a difference?

- or compare completers to all data models?

which is best?

glmmTMB(sae ~ time:group + time + sex + age + (1 | site/group),
data,
binomial,
na.omit)



# C

Infection time - Enrollment time

censor: lastFU whether it is larger than 12 months

```{r}
#loading data
surv_dat <- read_xlsx("Project 4/Q2c.xlsx",
                      col_types = c("numeric", "numeric", "numeric", "numeric", "logical"))
```

```{r}
library(survival)

surv_obj <- Surv(time = surv_dat$LastFUTime,
                 event = surv_dat$Infection,
                # type = 'right',
                 origin = surv_dat$EnrollmentTime)


km_fit <- survfit(surv_obj ~ 1, data = surv_dat)

plot(km_fit, xlab = 'Time to Infection (days)', ylab = 'Survival Probability')
```


```{r}
#calculating the prob of infection within 12 months:
summary(km_fit, times = 365) #prob = 0.708

#1-prob @ 365
1 - 0.708 # 0.292

#lower of prob:
summary(km_fit, times = 365)$lower
#0.685

#upper of prob:
summary(km_fit, times = 365)$upper
#0.732

#conf int:

#lower:
1 - 0.732
#0.268

#upper:
1 - 0.685
#0.315

# 0.292 (0.268, 0.315)
```

```{r}
#median time to infection
summary(km_fit)$table['median']
#doesn't reach halfway point
```

```{r}
#mean time to infection
summary(km_fit)$table

summary(km_fit)$table['rmean']
```


```{r}
#fitting on parametric models to find median time to infection

#exponential distribution:
exp_fit <- survreg(surv_obj ~ 1, data = surv_dat,
                   dist = 'exponential')
#median time to infection: 974.173


#weibull distribution:
weibull_fit <- survreg(surv_obj ~ 1, data = surv_dat,
                       dist =  'weibull')
#median time to infection: 453.668

#log-normal distribution:
logn_fit <- survreg(surv_obj ~ 1, data = surv_dat,
                    dist = 'lognormal')
#median time to infection: 457.345

#to find median time to infection:
predict(logn_fit, type = 'quantile', p = 0.5)
```


```{r, use for plotting parametric mods!}
exp_fit2 <- flexsurvreg(surv_obj ~ 1, data = surv_dat,
                        dist = 'exponential')

weibull_fit2 <- flexsurvreg(surv_obj ~ 1, data = surv_dat,
                        dist = 'weibull')

logn_fit2 <- flexsurvreg(surv_obj ~ 1, data = surv_dat,
                        dist = 'lognormal')


#median time to infection:
summary(exp_fit2, type = 'quantile', quantiles = 0.5) #same as before
```



```{r}
#plotting parametric + km models:
library(flexsurv)
library(survminer)

plot(km_fit, xlab = 'Time to Infection (days)', ylab = 'Survival Probability',
     conf.int = FALSE, lwd = 2)


#time grid:
time_grid <- seq(0, max(surv_dat$LastFUTime, na.rm = TRUE), by = 1)

plot(exp_fit2, col = 'blue', add = TRUE, t = time_grid, lwd = 2, ci = FALSE)
plot(weibull_fit2, col = 'red', add = TRUE, t = time_grid, lwd = 2, ci = FALSE)
plot(logn_fit2, col = 'green', add = TRUE, t = time_grid, lwd = 2, ci = FALSE)

legend('bottomleft', legend = c('KM (nonparametric)', 'Exponential', 'Weibull', 'Log-Normal'),
       col = c('black', 'blue', 'red', 'green'),
       lty = c(1, 2, 3, 4, 5), lwd = 2, bty = 'n')
```

```{r, eval=FALSE}
#exp survival fun:
#S(t) = exp(-lambda*t), lambda = exp(rate param)

#rate param:
exp_lambda <- exp_fit$coefficients[1]
#scale param:
exp_scale <- exp(exp_lambda)
#exp survival probs:
exp_surv <- exp(-surv_dat$LastFUTime/exp_scale)

#exp line:
lines(surv_dat$LastFUTime, exp_surv, col = 'blue', lwd = 2)




#exp line:

exp_pred <- predict(exp_fit, type = 'response', newdata = data.frame(time = surv_dat$LastFUTime))
#survival function:
exp_surv <- exp(-exp_pred)
lines(surv_dat$LastFUTime, exp_surv, col = 'blue')


#weibull line:

weibull_pred <- predict(weibull_fit, type = 'response', newdata = data.frame(time = surv_dat$LastFUTime))
weibull_surv <- exp(-weibull_pred)
lines(surv_dat$LastFUTime, weibull_surv, col = 'red')


  
  

#time grid:
time_grid <- seq(0, max(surv_dat$LastFUTime, na.rm = TRUE), by = 1)

#extracting survival preds from models:
exp_pred <- predict(exp_fit, type = 'survival', t = time_grid, newdata = data.frame(1))
weibull_pred <- summary(weibull_fit, t = time_grid)
logn_pred <- summary(logn_fit, t = time_grid)

#extracting survival estimates:
lines(time_grid, exp_pred, col = 'blue', ci = FALSE)
lines(time_grid, weibull_pred[[1]]$est, col = 'red', ci = FALSE)
lines(logn_fit, col = 'green', ci = FALSE)
```


```{r}
#spline based model:
spline_fit <- stpm2(Surv(EnrollmentTime, LastFUTime, Infection) ~ 1, data = surv_dat, df = 4)

plot(km_fit, xlab = 'Time to Infection (days)', ylab = 'Survival Probability',
     conf.int = FALSE, lwd = 2)

plot(spline_fit, col = 'blue', add = TRUE, t = time_grid, lwd = 2)
```

