---
title: "Project 1 - Longitudinal"
author: "Lupe Antonio"
date: "2025-02-23"
output: html_document
---

```{r setup, include=FALS, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(gtsummary)
library(corrplot)
library(tidymodels)
library(gridExtra)
library(GGally)

library(gee)
library(lme4)
library(nlme)
```


## loading & cleaning data
```{r,message=FALSE}
#loading data
data <- read_csv("data.csv")

dat <- data %>%
  mutate(day = as.factor(day),
         Gender = as.factor(Gender),
         Treatment_Group = as.factor(Treatment_Group))
```


## notes:

non-monotone missingness

missing at random?

GEE vs LMM vs GLM

marginal vs individual effect over time

binary vs continuous

(potentially want to verify if it is missing at random? using the if at least one of their time points is missing, then we can see that most of those if not all are the ones with greater memory score overall)


## eda

```{r}
head(dat)
```


## summary stats tbl using `gtsummary` 
```{r}
grp_sum <- dat %>%
  group_by(Treatment_Group) %>%
  summarise(unique_subjects = n_distinct(Subject_ID))

gender_sum <- dat %>%
  group_by(Gender) %>%
  summarise(unique_subjects = n_distinct(Subject_ID))

trt_day <- dat %>%
  group_by(Treatment_Group, day) %>%
  summarise(unique_subjects = n_distinct(Subject_ID),
            N = n())


#using unique subjects, finding the stats
unique_sum <- dat %>%
  distinct(Subject_ID, Age, Gender, Treatment_Group) %>%
  group_by(Treatment_Group) %>%
  summarise(avg_age = mean(Age),
            age_range = paste(min(Age),',',max(Age)),
            female_ct = sum(Gender == 'F'),
            male_ct = sum(Gender == 'M'),
            N = n())
```

```{r, table 1}
sum_tble <- dat %>%
  select(-Subject_ID, -MEM_comp) %>%
  mutate(
    Gender = case_when(Gender == 'M' ~ 'Male',
                       Gender == 'F' ~ 'Female'),
    
    Treatment_Group = case_when(Treatment_Group == 'A' ~ 'Placebo',
                                Treatment_Group == 'B' ~ 'Low dose',
                                Treatment_Group == 'C' ~ 'High dose')) %>%
    
   # day = case_when(day == '0' ~ 'Day 0',
   #                 day == '5' ~ 'Day 5',
   #                 day == '19' ~ 'Day 19',
   #                 day == '90' ~ 'Day 90')) %>%
  
  tbl_summary(by = Treatment_Group,
              statistic = list(all_continuous() ~ '{mean}({min}, {max})',
                               all_categorical() ~ '{n} ({p}%)'),
              digits = all_continuous() ~ 3,
              label = list(day ~ 'Day')) %>%
  modify_spanning_header(c('stat_1', 'stat_2', 'stat_3') ~ '**Treatment Group**') %>%
  modify_header(label = '**Variable**',
                stat_1 = paste('**High dose**, N = ', grp_sum$unique_subjects[3]),
                stat_2 = paste('**Low dose**, N = ', grp_sum$unique_subjects[2]),
                stat_3 = paste('**Placebo**, N = ', grp_sum$unique_subjects[1])) %>%
  bold_labels() %>%
  modify_caption('Table 1. Summary statistics')

#or patient characteristics
```


#modifying table vals
```{r}
sum_tble %>%
    modify_table_body(~ .x %>%
                      #editing Age vals
                      mutate(stat_1 = ifelse(variable == 'Age', paste(unique_sum$avg_age[3],
                                                                      '(', unique_sum$age_range[3], ')'),
                                             stat_1),
                             stat_2 = ifelse(variable == 'Age', paste(round(unique_sum$avg_age[2], 3),
                                                                      '(', unique_sum$age_range[2], ')'),
                                             stat_2),
                             stat_3 = ifelse(variable == 'Age', paste(round(unique_sum$avg_age[1], 3),
                                                                      '(', unique_sum$age_range[1], ')'),
                                             stat_3),
                             
                             #editing Gender = F vals
                             stat_1 = ifelse(variable == 'Gender' & label == 'Female',
                                             paste(unique_sum$female_ct[3], '(',
                                                   round((unique_sum$female_ct[3]/unique_sum$N[3])*100, 2), '%)'),
                                             stat_1),
                             stat_2 = ifelse(variable == 'Gender' & label == 'Female',
                                             paste(unique_sum$female_ct[2], '(',
                                                   round((unique_sum$female_ct[2]/unique_sum$N[2])*100, 2), '%)'),
                                             stat_2),
                             stat_3 = ifelse(variable == 'Gender' & label == 'Female',
                                             paste(unique_sum$female_ct[1], '(',
                                                   round((unique_sum$female_ct[1]/unique_sum$N[1])*100, 2), '%)'),
                                             stat_3),
                             #editing Gender = M vals
                             stat_1 = ifelse(variable == 'Gender' & label == 'Male',
                                             paste(unique_sum$male_ct[3], '(',
                                                   round((unique_sum$male_ct[3]/unique_sum$N[3])*100, 2), '%)'),
                                             stat_1),
                             stat_2 = ifelse(variable == 'Gender' & label == 'Male',
                                             paste(unique_sum$male_ct[2], '(',
                                                   round((unique_sum$male_ct[2]/unique_sum$N[2])*100, 2), '%)'),
                                             stat_2),
                             stat_3 = ifelse(variable == 'Gender' & label == 'Male',
                                             paste(unique_sum$male_ct[1], '(',
                                                   round((unique_sum$male_ct[1]/unique_sum$N[1])*100, 2), '%)'),
                                             stat_3),
                             #editing Day vals bc percentages:
                             #day 0
                             stat_1 = ifelse(variable == 'day' & label == '0',
                                             paste(trt_day$unique_subjects[9], '(',
                                                   round((trt_day$unique_subjects[9]/trt_day$N[9])*100,
                                                         2), '%)'),
                                             stat_1),
                             stat_2 = ifelse(variable == 'day' & label == '0',
                                             paste(trt_day$unique_subjects[5], '(',
                                                   round((trt_day$unique_subjects[5]/trt_day$N[5])*100,
                                                         2), '%)'),
                                             stat_2),
                             stat_3 = ifelse(variable == 'day' & label == '0',
                                             paste(trt_day$unique_subjects[1], '(',
                                                   round((trt_day$unique_subjects[1]/trt_day$N[1])*100,
                                                         2), '%)'),
                                             stat_3),
                             #day 5
                             stat_1 = ifelse(variable == 'day' & label == '5',
                                             paste(trt_day$unique_subjects[10], '(',
                                                   round((trt_day$unique_subjects[10]/trt_day$N[9])*100,
                                                         2), '%)'),
                                             stat_1),
                             stat_2 = ifelse(variable == 'day' & label == '5',
                                             paste(trt_day$unique_subjects[6], '(',
                                                   round((trt_day$unique_subjects[6]/trt_day$N[5])*100,
                                                         2), '%)'),
                                             stat_2),
                             stat_3 = ifelse(variable == 'day' & label == '5',
                                             paste(trt_day$unique_subjects[2], '(',
                                                   round((trt_day$unique_subjects[2]/trt_day$N[1])*100,
                                                         2), '%)'),
                                             stat_3),
                             #day 19
                             stat_1 = ifelse(variable == 'day' & label == '19',
                                             paste(trt_day$unique_subjects[11], '(',
                                                   round((trt_day$unique_subjects[11]/trt_day$N[9])*100,
                                                         2), '%)'),
                                             stat_1),
                             stat_2 = ifelse(variable == 'day' & label == '19',
                                             paste(trt_day$unique_subjects[7], '(',
                                                   round((trt_day$unique_subjects[7]/trt_day$N[5])*100,
                                                         2), '%)'),
                                             stat_2),
                             stat_3 = ifelse(variable == 'day' & label == '19',
                                             paste(trt_day$unique_subjects[3], '(',
                                                   round((trt_day$unique_subjects[3]/trt_day$N[1])*100,
                                                         2), '%)'),
                                             stat_3),
                             #day 90
                             stat_1 = ifelse(variable == 'day' & label == '90',
                                             paste(trt_day$unique_subjects[12], '(',
                                                   round((trt_day$unique_subjects[12]/trt_day$N[9])*100,
                                                         2), '%)'),
                                             stat_1),
                             stat_2 = ifelse(variable == 'day' & label == '90',
                                             paste(trt_day$unique_subjects[8], '(',
                                                   round((trt_day$unique_subjects[8]/trt_day$N[5])*100,
                                                         2), '%)'),
                                             stat_2),
                             stat_3 = ifelse(variable == 'day' & label == '90',
                                             paste(trt_day$unique_subjects[4], '(',
                                                   round((trt_day$unique_subjects[4]/trt_day$N[1])*100,
                                                         2), '%)'),
                                             stat_3)))
```


```{r}
dat %>%
  group_by(Treatment_Group, day) %>%
  summarise(subj_ct = n_distinct(Subject_ID))
```




```{r}
dat %>%
  filter(Treatment_Group == 'A') %>%
  unique()


#using unique subjects, finding the stats
dat %>%
  distinct(Subject_ID, Age, Gender, Treatment_Group) %>%
  group_by(Treatment_Group) %>%
  summarise(avg_age = mean(Age),
            age_range = paste(min(Age),',',max(Age)),
            female_ct = sum(Gender == 'F'),
            male_ct = sum(Gender == 'M'))
```


##table or plot for showcasing memory scores for each trt grp by day. 
```{r}
dat %>%
  select(-Subject_ID) %>%
  mutate(
    Gender = case_when(Gender == 'M' ~ 'Male',
                       Gender == 'F' ~ 'Female'),
    
    Treatment_Group = case_when(Treatment_Group == 'A' ~ 'Placebo',
                                Treatment_Group == 'B' ~ 'Low dose',
                                Treatment_Group == 'C' ~ 'High dose')) %>%
  tbl_summary(by = day,
              statistic = list(all_continuous() ~ '{mean}({sd})',
                               all_categorical() ~ '{n} ({p}%)'),
              digits = all_continuous() ~ 3,
              label = list(day ~ 'Day')) %>%
  modify_spanning_header(c('stat_1', 'stat_2', 'stat_3') ~ '**Day**') %>%
  modify_header(label = '**Variable**') %>%
  bold_labels() %>%
  modify_caption('Table 1. Summary statistics')
```



## correlation plots
```{r}
#converting covariates to numeric
cor_rec <- model.matrix(MEM_comp ~ ., data = dat[,2:6])

#adding back into matrix
cor_rec[, 1] = dat$MEM_comp

#plot
corrplot(cor(cor_rec), method = 'circle', type = 'full')
```


## qq-plot for response
```{r, message=FALSE}
qqnorm(dat$MEM_comp)
qqline(dat$MEM_comp, col = 'steelblue')
```


## histogram for response
```{r}
ggplot(data = dat, aes(x = MEM_comp)) +
  geom_histogram()
```

## with sqrt transformation
```{r}
ggplot(data = dat, aes(x = sqrt(MEM_comp))) +
  geom_histogram(bins = 25)
```


## fitting model(DONT FORGET AGE & GENDER):

$$
Y_{ij} = \beta_0 + \beta_{1}^{(1)} t_j + \beta_{1}^{(2)} t_j + \beta_{1}^{(3)} t_j + \beta_2 d_1 + \beta_2 d_2 +  \\
         \beta_4^{(1)}t_j \cdot d_1 + \beta_4^{(2)}t_j \cdot d_1 + \beta_4^{(3)}t_j \cdot d_1 + \\
         \beta_5^{(1)}t_j \cdot d_1 + \beta_5^{(2)}t_j \cdot d_1 + \beta_5^{(3)}t_j \cdot d_1 + e_{ij}
$$

Note:

- $\beta_1$ has three different dummy variables. They correspond to the three different time points, where Day 0 is the baseline. Thus, superscript 1 corresponds to Day 5, superscript 2 to Day 19, and superscript 3 to Day 90.

- $\beta_2d_1$ is the coefficient for those in the Low Dose group, where d_1 takes in values either 0 or 1 since it is a categorical/binary variable.

- $\beta_2d_2$ is the coefficient for those in the High Dose group, where d_2 takes in values either 0 or 1 since it is a categorical/binary variable.

- $\beta_4$ also has three dummy variables. They correspond to the three different time points and whether they were in the Low Dose group or not. Thus, superscript 1 correspond to those at Day 5 that were in the Low Dose group, superscript 2 for those at Day 19 that were in the Low Dose group, and superscript 3 for those at Day 90 that were in the Low Dose group.

- $\beta_4$ also has three dummy variables. They correspond to the three different time points and whether they were in the High Dose group or not. Thus, superscript 1 correspond to those at Day 5 that were in the High Dose group, superscript 2 for those at Day 19 that were in the High Dose group, and superscript 3 for those at Day 90 that were in the High Dose group.



#generalized linear mixed effects model
```{r, glm}
mod <- glmer(MEM_comp ~ day + Age + Gender + Treatment_Group + (1 | Subject_ID),
             data = dat, family = gaussian,
             nAGQ = 0)

summary(mod)
```


#linear mixed effects model: subj are random effects, fixed effects are other covariates
```{r}
mod <- lmer() #used for multiple random effects, downside is can't specify corr structure. 
```


#using lme() to specify the correlation matrix: Unstructured. 
```{r}
mod <- lme(MEM_comp ~ day + Treatment_Group + day*Treatment_Group, random = ~ 1 | Subject_ID,
           data = dat, method = 'REML', correlation = corSymm())

summary(mod)
```

#varcov matrix
```{r}
intercept_var <- VarCorr(mod)[1,1]
residual_var <- VarCorr(mod)[2,1]

#covariance matrix for subject Y_i
matrix(rep(as.numeric(intercept_var), 14^2), nrow = 14) +
  diag(residual_var, nrow = 14)
```


#fitting model w fixed effects too (i.e. age & gender)
#this is only with random intercepts.
```{r}
mod2 <- lme(MEM_comp ~ day + Age + Gender + Treatment_Group + day*Treatment_Group,
            random = ~ 1 | Subject_ID,
            data = dat, method = 'REML', correlation = corSymm())

summary(mod2)
```

#fitting model with fixed effects (Age & Gender) + including random intercept & slope (for Day)
#random slop isn't working...
```{r}
mod3 <- lme(MEM_comp ~ day + Age + Gender + Treatment_Group + day*Treatment_Group,
            random = ~ day | Subject_ID,
            data = dat, method = 'REML', correlation = corSymm(),
            control = lmeControl(opt = 'optim', maxIter = 20000000))

summary(mod3)
```




#interpretations for mod with random intercept only:

- `Intercept`: the overall average MEM_comp score for Females at Day 0 and in the Placebo group is 0.8960.

- `day5`: the MEM_comp score for an individual at Day 5, holding all other predictors constant, decreases by 0.0951 compared to when at Day 0. 

- `day19`: the MEM_comp score for an individual at Day 19, holding all other predictors constant, decreases by 0.2330 compared to when at Day 0.

- `day90`: the MEM_comp score for an individual at Day 90, holding all other predictors constant, decreases by 0.1495 compared to when at Day 0.

- `Age`: the MEM_comp score for an individual decreases by 0.0207 as age increase by 1 year.

- `GenderM`: the MEM_comp score for an Male individual, holding all other predictors constant, increases by 0.0158 compared to a Female individual.

- `Treatment_GroupB`: the MEM_comp score for an individual in the Low Dose treatment group, holding all other predictors constant, decreases by 0.0917 commpared to those in the Placebo group.

- `Treatment_GroupC`: the MEM_comp score for an individual in the High Dose treatment group, holding all other predictors constant, decreases by 0.1881 commpared to those in the Placebo group.

- `day5:Treatment_GroupB`: the MEM_comp score for an individual at Day 5 in the Low Dose group, holding all other predictors constant, decreases by 0.0515 compared to one at Day 0 in the Placebo group.

- `day19:Treatment_GroupB`: the MEM_comp score for an individual at Day 19 in the Low Dose group, holding all other predictors constant, increases by 0.1749 compared to one at Day 0 in the Placebo group.

- `day90:Treatment_GroupB`: the MEM_comp score for an individual at Day 90 in the Low Dose group, holding all other predictors constant, increases by 0.2544 compared to one at Day 0 in the Placebo group.

- `day5:Treatment_GroupC`: the MEM_comp score for an individual at Day 5 in the High Dose group, holding all other predictors constant, increases by 0.3860 compared to one at Day 0 in the Placebo group.

- `day19:Treatment_GroupC`: the MEM_comp score for an individual at Day 19 in the High Dose group, holding all other predictors constant, increases by 0.6547 compared to one at Day 0 in the Placebo group.

- - `day90:Treatment_GroupC`: the MEM_comp score for an individual at Day 90 in the High Dose group, holding all other predictors constant, increases by 0.2993 compared to one at Day 0 in the Placebo group.


- `random intercept`: Given the average MEM_comp score at baseline, each individual might deviate from this baseline intercept by 0.8401421. 



^ original model (random intercept model? def mixed effects)





## model with only complete cases:

there are 47 total unique subjects, so complete_days should have 188 possible combos.

```{r, complete cases data}
#creating all possible combinations between subj and day
#complete_days <- expand.grid(Subject_ID = unique(dat$Subject_ID),
#                             day = unique(dat$day))

#now joining with data to create complete data
#contains missingness
#all_data <- complete_days %>%
#  left_join(dat, by = c('Subject_ID', 'day'))
  

#data containing only subjects with complete data
#complete_data <- all_data %>%
#  group_by(Subject_ID) %>%
#  filter(all(!is.na(day))) %>%
#  ungroup()



#counting all time points each subj has
subject_days <- dat %>%
  group_by(Subject_ID) %>%
  summarise(day_ct = n_distinct(day))

#only subjects with all days
complete_subj <- subject_days %>%
  filter(day_ct == '4') %>%
  pull(Subject_ID) #only 24 subj's

#complete data; only subjects with all days
complete_data <- dat %>%
  filter(Subject_ID %in% complete_subj) #should be only 96 entries (24*4)
```


```{r, model w complete data}
complete_mod <- lme(MEM_comp ~ day + Age + Gender + Treatment_Group + day*Treatment_Group,
                    random = ~ 1 | Subject_ID,
                    data = dat, method = 'REML', correlation = corSymm())

summary(complete_mod)
```







