---
title: "Project 1 - Longitudinal"
author: "Lupe Antonio"
date: "2025-02-23"
output: html_document
---

```{r setup, include=FALS, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(gtsummary)
library(corrplot)
library(tidymodels)
library(gridExtra)
library(GGally)
```


## loading & cleaning data
```{r,message=FALSE}
#loading data
data <- read_csv("Project 1 - Longitudinal/data.csv")

dat <- data %>%
  mutate(day = as.factor(day),
         Gender = as.factor(Gender),
         Treatment_Group = as.factor(Treatment_Group))
```


## notes:

non-monotone missingness

missing at random?

GEE vs LMM vs GLM

marginal vs individual effect over time

binary vs continuous

(potentially want to verify if it is missing at random? using the if at least one of their time points is missing, then we can see that most of those if not all are the ones with greater memory score overall)


## eda

```{r}
head(dat)
```


## summary stats tbl using `gtsummary` 
```{r}
grp_sum <- dat %>%
  group_by(Treatment_Group) %>%
  summarise(unique_subjects = n_distinct(Subject_ID))

gender_sum <- dat %>%
  group_by(Gender) %>%
  summarise(unique_subjects = n_distinct(Subject_ID))

trt_day <- dat %>%
  group_by(Treatment_Group, day) %>%
  summarise(unique_subjects = n_distinct(Subject_ID),
            N = n())


#using unique subjects, finding the stats
unique_sum <- dat %>%
  distinct(Subject_ID, Age, Gender, Treatment_Group) %>%
  group_by(Treatment_Group) %>%
  summarise(avg_age = mean(Age),
            age_range = paste(min(Age),',',max(Age)),
            female_ct = sum(Gender == 'F'),
            male_ct = sum(Gender == 'M'),
            N = n())
```

```{r, table 1}
sum_tble <- dat %>%
  select(-Subject_ID, -MEM_comp) %>%
  mutate(
    Gender = case_when(Gender == 'M' ~ 'Male',
                       Gender == 'F' ~ 'Female'),
    
    Treatment_Group = case_when(Treatment_Group == 'A' ~ 'Placebo',
                                Treatment_Group == 'B' ~ 'Low dose',
                                Treatment_Group == 'C' ~ 'High dose')) %>%
    
   # day = case_when(day == '0' ~ 'Day 0',
   #                 day == '5' ~ 'Day 5',
   #                 day == '19' ~ 'Day 19',
   #                 day == '90' ~ 'Day 90')) %>%
  
  tbl_summary(by = Treatment_Group,
              statistic = list(all_continuous() ~ '{mean}({min}, {max})',
                               all_categorical() ~ '{n} ({p}%)'),
              digits = all_continuous() ~ 3,
              label = list(day ~ 'Day')) %>%
  modify_spanning_header(c('stat_1', 'stat_2', 'stat_3') ~ '**Treatment Group**') %>%
  modify_header(label = '**Variable**',
                stat_1 = paste('**High dose**, N = ', grp_sum$unique_subjects[3]),
                stat_2 = paste('**Low dose**, N = ', grp_sum$unique_subjects[2]),
                stat_3 = paste('**Placebo**, N = ', grp_sum$unique_subjects[1])) %>%
  bold_labels() %>%
  modify_caption('Table 1. Summary statistics')

#or patient characteristics
```


#modifying table vals
```{r}
sum_tble %>%
    modify_table_body(~ .x %>%
                      #editing Age vals
                      mutate(stat_1 = ifelse(variable == 'Age', paste(unique_sum$avg_age[3],
                                                                      '(', unique_sum$age_range[3], ')'),
                                             stat_1),
                             stat_2 = ifelse(variable == 'Age', paste(round(unique_sum$avg_age[2], 3),
                                                                      '(', unique_sum$age_range[2], ')'),
                                             stat_2),
                             stat_3 = ifelse(variable == 'Age', paste(round(unique_sum$avg_age[1], 3),
                                                                      '(', unique_sum$age_range[1], ')'),
                                             stat_3),
                             
                             #editing Gender = F vals
                             stat_1 = ifelse(variable == 'Gender' & label == 'Female',
                                             paste(unique_sum$female_ct[3], '(',
                                                   round((unique_sum$female_ct[3]/unique_sum$N[3])*100, 2), '%)'),
                                             stat_1),
                             stat_2 = ifelse(variable == 'Gender' & label == 'Female',
                                             paste(unique_sum$female_ct[2], '(',
                                                   round((unique_sum$female_ct[2]/unique_sum$N[2])*100, 2), '%)'),
                                             stat_2),
                             stat_3 = ifelse(variable == 'Gender' & label == 'Female',
                                             paste(unique_sum$female_ct[1], '(',
                                                   round((unique_sum$female_ct[1]/unique_sum$N[1])*100, 2), '%)'),
                                             stat_3),
                             #editing Gender = M vals
                             stat_1 = ifelse(variable == 'Gender' & label == 'Male',
                                             paste(unique_sum$male_ct[3], '(',
                                                   round((unique_sum$male_ct[3]/unique_sum$N[3])*100, 2), '%)'),
                                             stat_1),
                             stat_2 = ifelse(variable == 'Gender' & label == 'Male',
                                             paste(unique_sum$male_ct[2], '(',
                                                   round((unique_sum$male_ct[2]/unique_sum$N[2])*100, 2), '%)'),
                                             stat_2),
                             stat_3 = ifelse(variable == 'Gender' & label == 'Male',
                                             paste(unique_sum$male_ct[1], '(',
                                                   round((unique_sum$male_ct[1]/unique_sum$N[1])*100, 2), '%)'),
                                             stat_3),
                             #editing Day vals bc percentages:
                             #day 0
                             stat_1 = ifelse(variable == 'day' & label == '0',
                                             paste(trt_day$unique_subjects[9], '(',
                                                   round((trt_day$unique_subjects[9]/trt_day$N[9])*100,
                                                         2), '%)'),
                                             stat_1),
                             stat_2 = ifelse(variable == 'day' & label == '0',
                                             paste(trt_day$unique_subjects[5], '(',
                                                   round((trt_day$unique_subjects[5]/trt_day$N[5])*100,
                                                         2), '%)'),
                                             stat_2),
                             stat_3 = ifelse(variable == 'day' & label == '0',
                                             paste(trt_day$unique_subjects[1], '(',
                                                   round((trt_day$unique_subjects[1]/trt_day$N[1])*100,
                                                         2), '%)'),
                                             stat_3),
                             #day 5
                             stat_1 = ifelse(variable == 'day' & label == '5',
                                             paste(trt_day$unique_subjects[10], '(',
                                                   round((trt_day$unique_subjects[10]/trt_day$N[9])*100,
                                                         2), '%)'),
                                             stat_1),
                             stat_2 = ifelse(variable == 'day' & label == '5',
                                             paste(trt_day$unique_subjects[6], '(',
                                                   round((trt_day$unique_subjects[6]/trt_day$N[5])*100,
                                                         2), '%)'),
                                             stat_2),
                             stat_3 = ifelse(variable == 'day' & label == '5',
                                             paste(trt_day$unique_subjects[2], '(',
                                                   round((trt_day$unique_subjects[2]/trt_day$N[1])*100,
                                                         2), '%)'),
                                             stat_3),
                             #day 19
                             stat_1 = ifelse(variable == 'day' & label == '19',
                                             paste(trt_day$unique_subjects[11], '(',
                                                   round((trt_day$unique_subjects[11]/trt_day$N[9])*100,
                                                         2), '%)'),
                                             stat_1),
                             stat_2 = ifelse(variable == 'day' & label == '19',
                                             paste(trt_day$unique_subjects[7], '(',
                                                   round((trt_day$unique_subjects[7]/trt_day$N[5])*100,
                                                         2), '%)'),
                                             stat_2),
                             stat_3 = ifelse(variable == 'day' & label == '19',
                                             paste(trt_day$unique_subjects[3], '(',
                                                   round((trt_day$unique_subjects[3]/trt_day$N[1])*100,
                                                         2), '%)'),
                                             stat_3),
                             #day 90
                             stat_1 = ifelse(variable == 'day' & label == '90',
                                             paste(trt_day$unique_subjects[12], '(',
                                                   round((trt_day$unique_subjects[12]/trt_day$N[9])*100,
                                                         2), '%)'),
                                             stat_1),
                             stat_2 = ifelse(variable == 'day' & label == '90',
                                             paste(trt_day$unique_subjects[8], '(',
                                                   round((trt_day$unique_subjects[8]/trt_day$N[5])*100,
                                                         2), '%)'),
                                             stat_2),
                             stat_3 = ifelse(variable == 'day' & label == '90',
                                             paste(trt_day$unique_subjects[4], '(',
                                                   round((trt_day$unique_subjects[4]/trt_day$N[1])*100,
                                                         2), '%)'),
                                             stat_3)))
```


```{r}
dat %>%
  group_by(Treatment_Group, day) %>%
  summarise(subj_ct = n_distinct(Subject_ID))
```




```{r}
dat %>%
  filter(Treatment_Group == 'A') %>%
  unique()


#using unique subjects, finding the stats
dat %>%
  distinct(Subject_ID, Age, Gender, Treatment_Group) %>%
  group_by(Treatment_Group) %>%
  summarise(avg_age = mean(Age),
            age_range = paste(min(Age),',',max(Age)),
            female_ct = sum(Gender == 'F'),
            male_ct = sum(Gender == 'M'))
```


##table or plot for showcasing memory scores for each trt grp by day. 
```{r}
dat %>%
  select(-Subject_ID) %>%
  mutate(
    Gender = case_when(Gender == 'M' ~ 'Male',
                       Gender == 'F' ~ 'Female'),
    
    Treatment_Group = case_when(Treatment_Group == 'A' ~ 'Placebo',
                                Treatment_Group == 'B' ~ 'Low dose',
                                Treatment_Group == 'C' ~ 'High dose')) %>%
  tbl_summary(by = day,
              statistic = list(all_continuous() ~ '{mean}({sd})',
                               all_categorical() ~ '{n} ({p}%)'),
              digits = all_continuous() ~ 3,
              label = list(day ~ 'Day')) %>%
  modify_spanning_header(c('stat_1', 'stat_2', 'stat_3') ~ '**Day**') %>%
  modify_header(label = '**Variable**') %>%
  bold_labels() %>%
  modify_caption('Table 1. Summary statistics')
```



## correlation plots
```{r}
#converting covariates to numeric
cor_rec <- model.matrix(MEM_comp ~ ., data = dat[,2:6])

#adding back into matrix
cor_rec[, 1] = dat$MEM_comp

#plot
corrplot(cor(cor_rec), method = 'circle', type = 'full')
```


## qq-plot for response
```{r, message=FALSE}
qqnorm(dat$MEM_comp)
qqline(dat$MEM_comp, col = 'steelblue')
```


## histogram for response
```{r}
ggplot(data = dat, aes(x = MEM_comp)) +
  geom_histogram()
```

## with sqrt transformation
```{r}
ggplot(data = dat, aes(x = sqrt(MEM_comp))) +
  geom_histogram(bins = 25)
```







