---
title: "Longitudinal Analysis"
output: pdf_document
author: "Liner Ge"
date: "2025-02-23"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(tidyr)
library(tidyverse)
library(table1)
library(corrplot)
```

## Scientific Question

### Primary objective

- Determine whether the effect of low dose DAR-0100A and high dose DAR-0100A on improving memory as measured by the composite memory score are significantly higher than the effect of placebo at day 5, day 19, and day 90.

- Determine whether the treatment effects differ over time within placebo group, low dose group and high dose group.

- Determine whether the trajectory of treatment effects over time differ between placebo group, low dose group and high dose group.

### Sensitivity analysis

## EDA

### Table 1

```{r}
data = read.csv("./data.csv") %>% 
  mutate(day =case_when(day == 0 ~ "Day 0",
                        day == 5 ~ "Day 5",
                        day == 19 ~ "Day 19",
                        day == 90 ~ "Day 90"
                        ))

data_table1= data%>% 
  group_by(Subject_ID) %>% 
  filter(day == "Day 0")

table1(~ Age + factor(Gender) | Treatment_Group, data=data_table1)
```

### Spaghetti plot

```{r}
ggplot(data, aes(x=day, y = MEM_comp, group = Subject_ID))+
  geom_line()+ 
    stat_summary(aes(group = 1), geom = "point", fun.y = mean,
    shape = 16, size = 2, color = "red")+
  facet_grid(. ~ Treatment_Group)+
   labs(title="Change of composite memory score over time by treatment group",
       x="Time points", 
       y = "Composite memory score")+
  theme_light()
```


### Boxplot for displaying the composite memory score by group at different time points

```{r}
ggplot(data)+
  geom_boxplot(aes(x=day, y = MEM_comp, fill = Treatment_Group))+ 
  labs(title="Plot of composite memory score by treatment group at different time points",
       x="Time points", 
       y = "Composite memory score")+
  scale_fill_discrete(name = "Treatment group",
                      labels=c("A: Placebo","B: Low dose DAR-0100A","C: High dose DAR-0100A"))+
  theme_classic()
```

### Correlation matrix

```{r}
cor_data = data %>% 
  pivot_wider( names_from = day,
              values_from = MEM_comp) %>% 
  mutate(
    Age = as.numeric(Age),
    Gender = case_when(Gender == "M" ~ 1,
                            Gender == "F" ~ 0,
                            ),
         Treatment_Group =case_when(Treatment_Group=="A" ~ 0,
                                    Treatment_Group=="B" ~ 1,
                                    Treatment_Group=="C" ~ 2,)
         ) %>% 
  select(-Subject_ID)

corrplot(cor(cor_data, use='complete.obs'), addCoef.col = "red", type = "upper",tl.col="black", tl.srt=20)
```



